### 2011-10-31 doc/theses/vitasek/chapter-0-preface.tex
Thesis: in progress.

But the schema is not the only constraint that can be imposed on an XML document. Any textual or numerical value featured in the document may be subject to type constraints, such as the requirement to conform to a specific regular expression. Furthermore, the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies to schemas as well and will be the topic of this work. One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} well researched in \cite{sviro}.\\

From all the constraints that can be applied to an XML document by means of its schema, this work will focus on keys and foreign keys. Most important concepts in this field are introduced in \cite{keX} and formalized in the notions of \texttt{ID/IDREF/IDREFS} attributes in DTD and XSD and \texttt{xs:key/xs:keyref} structures in XSD (both in \cite{Bray:08:EML}).\\

The scope of this work is finally limited to the inference of \texttt{ID/IDREF/IDREFS} from existing XML documents. ID attributes were chosen over \texttt{xs:key} because the preliminary research found that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with ID attributes are much more common and it is much easier to obtain large data sets for experiments. % TODO something like even the search for ID sets is interesting enough?

@@

First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets, linear programming and mixed integer problem.

@@

Afterwards, we will show how to use an external MIP solver, demonstrate that this can take too long and how to use a heuristic approach to find good solutions much faster.

An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.

@@

\begin{algorithm}
\caption{Example Algorithm}
\label{listing-example}
\begin{algorithmic}
\REQUIRE $I$ input data
\REQUIRE $n$ maximum number of iterations
\ENSURE results found
\FOR{$i = 1 \to n$}
  \STATE \COMMENT{try to find a solution}
  \STATE $attempt \gets $ calculate possible solution from $I$
  \IF{$attempt$ is a valid solution}
    \RETURN $attempt$
  \ENDIF
  \RETURN "solution not found"
\ENDFOR
\end{algorithmic}
\end{algorithm}

>>>>

But the schema is not the only constraint that can be imposed on an XML document. Any textual or numerical value featured in the document may be subject to type constraints, such as the requirement to conform to a specific regular expression. Furthermore, the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies to schemas as well and will be the topic of this work. One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} researched in \cite{sviro}.\\

From all the constraints that can be applied to an XML document by means of its schema, this work will focus on keys and foreign keys. Most important concepts in this field are introduced in \cite{keX} and formalized in the notions of \texttt{ID}/\.\texttt{IDREF}/\.\texttt{IDREFS} attributes in DTD and XSD and \texttt{xs:key/xs:keyref} structures in XSD (both in \cite{Bray:08:EML}).\\

The scope of this work is finally limited to the inference of \texttt{ID}/\.\texttt{IDREF}/\.\texttt{IDREFS} from existing XML documents. \texttt{ID} attributes were chosen over \texttt{xs:key} because the preliminary research found that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with \texttt{ID} attributes are much more common and it is much easier to obtain large data sets for experiments. % TODO something like even the search for ID sets is interesting enough?

@@

First, we will introduce a few notions required throughout the work, such as XML tree, \texttt{ID} attributes, ID sets, linear programming and the mixed integer problem.

@@

Afterwards, we will show how to use an external MIP solver and demonstrate that this can take too much time. Next we show how to use a heuristic approach to find good solutions much faster.

An extension to jInfer for finding \texttt{ID} attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.

@@

\begin{algorithm}
\caption{Example Algorithm}
\label{listing-example}
\begin{algorithmic}
\REQUIRE $I$ input data
\REQUIRE $n$ maximum number of iterations
\ENSURE results found
\FOR{$i = 1 \to n$}
  \STATE \COMMENT{try to find a solution}
  \STATE $attempt \gets $ calculate possible solution from $I$
  \IF{$attempt$ is a valid solution}
    \RETURN $attempt$
  \ENDIF
  \RETURN ``solution not found''
\ENDFOR
\end{algorithmic}
\end{algorithm}

### 2011-10-31 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress.

We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data. This tree shall be called an \textit{XML tree}. For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the referenced article does.

@@

\begin{define}[Node equality]
	$v_1$ and $v_2$ are node equal, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
\end{define}

\begin{define}[Value equality]
	$v_1$ and $v_2$ are value equal, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
\end{define}

\section{ID, IDREF, IDREFS Attributes}
\label{section-definitions-id-attributes}

According to \cite{Bray:08:EML}, an XML attribute may have the type \textit{ID}, \textit{IDREF} or \textit{IDREFS} (among others). Following constraints are related to these types.

\begin{quote}
\textbf{Validity constraint: ID}

Values of type \textbf{ID} \textsc{must} match the Name production. A name \textsc{must not} appear more than once in an XML document as a value of this type; i.e., \textbf{ID} values \textsc{must} uniquely identify the elements which bear them.

\textbf{Validity constraint: One ID per Element Type}

An element type \textsc{must not} have more than one \textbf{ID} attribute specified.

\textbf{Validity constraint: ID Attribute Default}

An \textbf{ID} attribute \textsc{must} have a declared default of \textbf{\#IMPLIED} or \textbf{\#REQUIRED}.

\textbf{Validity constraint: IDREF}

Values of type \textbf{IDREF} \textsc{must} match the Name production, and values of type \textbf{IDREFS} \textsc{must} match Names; each Name \textsc{must} match the value of an \textbf{ID} attribute on some element in the XML document; i.e. \textbf{IDREF} values \textsc{must} match the value of some ID attribute.
\end{quote}

@@

%TODO

@@

%TODO compare them to ID attributes

@@

%TODO show how a XML key is defined using the keX notation % this is a terrible sentence

@@

Based on the requirements for an ID attribute from Section \ref{section-definitions-id-attributes} we will define ID set with the help of the following definition.

@@

That is, an ID set has images without repeating values and all the types are unique (an element cannot have more than one ID attribute).

@@

\subsection{IDREF and IDREFS Condition}

Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked IDREF.
\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i)\]
Furthermore if $m$ contains multivalued attributes, it is to be marked IDREFS.

>>>>

We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data. This tree shall be called an \textit{XML tree}. For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the cited article does.

@@

\begin{define}[Node equality]
	$v_1$ and $v_2$ are \textit{node equal}, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
\end{define}

\begin{define}[Value equality]
	$v_1$ and $v_2$ are \textit{value equal}, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
\end{define}

\section{\texttt{ID}, \texttt{IDREF}, \texttt{IDREFS} Attributes}
\label{section-definitions-id-attributes}

According to \cite{Bray:08:EML}, an XML attribute may have the type \texttt{ID}, \texttt{IDREF} or \texttt{IDREFS} (among others). Following constraints are related to these types.

\begin{quote}
\textbf{Validity constraint: \texttt{ID}}

Values of type \texttt{ID} \textsc{must} match the Name production. A name \textsc{must not} appear more than once in an XML document as a value of this type; i.e., \texttt{ID} values \textsc{must} uniquely identify the elements which bear them.

\textbf{Validity constraint: One \texttt{ID} per Element Type}

An element type \textsc{must not} have more than one \texttt{ID} attribute specified.

\textbf{Validity constraint: \texttt{ID} Attribute Default}

An \texttt{ID} attribute \textsc{must} have a declared default of \texttt{\#IMPLIED} or \texttt{\#REQUIRED}.

\textbf{Validity constraint: \texttt{IDREF}}

Values of type \texttt{IDREF} \textsc{must} match the Name production, and values of type \texttt{IDREFS} \textsc{must} match Names; each Name \textsc{must} match the value of an \texttt{ID} attribute on some element in the XML document; i.e. \texttt{IDREF} values \textsc{must} match the value of some \texttt{ID} attribute.
\end{quote}

@@

@@

% - compare them to ID attributes

@@

% - show how a XML key is defined using the keX notation % this is a terrible sentence

@@

Based on the requirements for an \texttt{ID} attribute from Section \ref{section-definitions-id-attributes} we will define ID set with the help of the following definition.

@@

That is, an ID set has images without repeating values and all the types are unique (an element cannot have more than one \texttt{ID} attribute).

@@

\subsection{\texttt{IDREF} and \texttt{IDREFS} Condition}

Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked \texttt{IDREF}.
\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i)\]
Furthermore if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.

### 2011-10-31 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \dataset{100-1000}. Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute''.

@@

Interestingly, the grammar extraction times tend to be shorter in the case when text data is not ignored, although this is inconclusive. However, there is a clear improvement of about 50 \% in case of model creation times. The conclusion then is to ignore the simple text node content whenever possible when finding ID attributes.

>>>>

Table \ref{table-experiments-fidax-and-hungry-idsets} lists the \texttt{ID} attributes found in both cases for this most extreme input, \dataset{100-1000}. Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as \texttt{ID} attribute''.

@@

Interestingly, the grammar extraction times tend to be shorter in the case when text data is not ignored, although this is inconclusive. However, there is a clear improvement of about 50 \% in case of model creation times. The conclusion then is to ignore the simple text node content whenever possible when finding \texttt{ID} attributes.
